---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- Use snippet 'render_markdown' for it -->

```{r,echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

# biblio

<!-- badges: start -->
[![cran checks](https://badges.cranchecks.info/worst/biblio.svg)](https://cran.r-project.org/web/checks/check_results_biblio.html)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/biblio)](https://cran.r-project.org/package=biblio)
![r-universe](https://kamapu.r-universe.dev/badges/biblio)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.10984056.svg)](https://doi.org/10.5281/zenodo.10984056)
<br>
[![R-CMD-check](https://github.com/kamapu/biblio/workflows/R-CMD-check/badge.svg)](https://github.com/kamapu/biblio/actions)
[![Codecov test coverage](https://codecov.io/gh/kamapu/biblio/branch/master/graph/badge.svg)](https://codecov.io/gh/kamapu/biblio?branch=master)
[![CRAN_downloads](http://cranlogs.r-pkg.org/badges/biblio)](https://cran.r-project.org/package=biblio)
[![total downloads](http://cranlogs.r-pkg.org/badges/grand-total/biblio)](https://cran.r-project.org/package=biblio)
<!-- badges: end -->


An R-package to manage bibliographic references. This package is focused on
data import-export to a data frame format based on the software
[**JabRef**](http://www.jabref.org/).

## Installing biblio

You can install this package from **GitHub**.

```{r, eval=FALSE}
library(remotes)
install_github("kamapu/biblio")
```

## Reading Bibtex Files

A copy of the references cited in the book of [Luebert & Pliscoff
(2016)](https://doi.org/10.5281/zenodo.60800) is also installed with this
package.

```{r}
library(biblio)
Bib <- read_bib(x = file.path(path.package("biblio"), "LuebertPliscoff.bib"))
Bib
```

## Scanning Cited References in R-Markdown Documents

An example of r-markdown document is also installed to test the
function `detect_keys()`, which is able to recognize the bibtexkeys in use
within an r-markdown document.

```{r}
my_document <- readLines(file.path(path.package("biblio"), "document.Rmd"))
cited_refs <- detect_keys(my_document)
cited_refs
```

Alternatively to the last option, you can use the function `read_rmd()` from
the package [`yamlme`](https://kamapu.github.io/rpkg/yamlme/).
Note that in this case the position of the citation may change because
`detect_keys()` will start counting lines after the yaml-header.

```{r}
library(yamlme)
my_document <- read_rmd(file.path(path.package("biblio"), "document.Rmd"))
cited_refs <- detect_keys(my_document)
cited_refs
```

The output shows the respective bibtexkey by its occurrence in the document
(line number).
With this output it is also possible to do some statistics.

```{r}
stats_refs <- aggregate(line ~ bibtexkey, data = cited_refs, FUN = length)

# Number of citations in text
sum(stats_refs$line)

# Number of cited articles
nrow(stats_refs)

# Respective frequency to know rare citations in text
stats_refs[order(stats_refs$line, decreasing = TRUE), ]
```

## Automatic Reference Lists

The package [`yamlme`](https://kamapu.github.io/rpkg/yamlme/) enables the use
of functions to write R-markdown documents and to render them.
In this context, the function `reflist()` will produce automatically a reference
list.

```{r eval = FALSE}
Bib2 <- subset(x = Bib, subset = as.integer(year) > 2005)
reflist(obj = Bib2, filename = "my_reflist")
```

Note that assigning the output of `reflist()` to a new object and using the
package `yamlme`, you may be able to modify the layout by using the function
`update()`. Remember to set 

```{r eval = FALSE}
library(yamlme)

# Subset and assignment
Bib2 <- subset(x = Bib, subset = as.integer(year) > 2005)
my_reflist <- reflist(x = Bib2, filename = "my_reflist", browse_file = FALSE)

# Updating and rendering
my_reflist <- update(
  object = my_reflist,
  title = "Reference List (PDF version)",
  author = "myself",
  output = "pdf_document",
  body = txt_body(
    "\\setlength{\\parindent}{-5mm}",
    "\\setlength{\\leftskip}{5mm}",
    "\\setlength{\\parskip}{8pt}"
  )
)
render_rmd(my_reflist, output_file = "my_reflist")
```

Note that rendering the updated document will also require the written bibtex
file, thus sharing an input bibtexfile or an `lib_df` object within an R-image
may be the most secure way to enable repeatability.
